---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# POMADE

<!-- badges: start -->
<!-- badges: end -->

The goal of POMADE package is to provide approximation and plot functions for conducting power analysis of the CHE, MLMA, and CE models for meta-analysis of dependent effect sizes developed by Vembye, Pustejovsky, & Pigott (2022). 

## Installation

You can install the development version of POMADE like so:

```{r, eval = FALSE}

# install.packages("devtools")
devtools::install_github("MikkelVembye/POMADE")

```


## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(POMADE)
## basic example code
```


# Reference

Vembye, Pustejovsky, & Pigott (2022). Power Approximations for Meta-Analysis of Dependent Effect sizes. <https://osf.io/preprints/metaarxiv/6tp9y/>
